---
title: "Oncoprint"
output: 
  html_document:
    df_print: paged
    theme: flatly
    number_sections: yes
    toc: yes
    toc_float: yes
    code_folding: hide
---

```{r}
library(tidyverse)
library(ComplexHeatmap)
library(msigdbr)
```

```{r}
PATH_DATA = "/lustre06/project/6004736/alvann/from_narval/220318_EZHIP/WES/clinical_seq_patients/out/02_oncoprint/pcgr.tiers.all.samples.tsv"
PATH_META_SAMPLE = "/lustre06/project/6004736/alvann/from_narval/220318_EZHIP/files/sample_meta.tsv"
PATH_META_PATIENT = "/lustre06/project/6004736/alvann/from_narval/220318_EZHIP/files/patient_meta.tsv"
```

```{r}
pcgr = data.table::fread(PATH_DATA) %>%
       filter(grepl('_RE1', VCF_SAMPLE_ID) | grepl('_T', VCF_SAMPLE_ID)) %>%
       separate(VCF_SAMPLE_ID, into = 'VCF_SAMPLE_ID', sep = '_')

pcgr 
```

```{r}
meta_pat = data.table::fread(PATH_META_PATIENT) %>% arrange(Patient_N)
meta_pat
```

```{r}
meta_samp = data.table::fread(PATH_META_SAMPLE)
```

```{r}
meta = full_join(meta_samp, meta_pat, by = 'Patient_N') %>%
       mutate(Response = tolower(Response), 
              Primary_location = tolower(Primary_location)) %>%
       filter(ID_triceps_signature_profyle %in% unique(pcgr$VCF_SAMPLE_ID))
```

# All TIERS 

## Count by gene 

```{r}
pcgr %>%
  filter(grepl('TIER', pcgr$TIER)) %>%
  count(SYMBOL, sort = T)
```

## Count by gene and genomic change 

```{r}
pcgr %>% 
  filter(grepl('TIER', pcgr$TIER)) %>%
  count(SYMBOL, PROTEIN_CHANGE, sort = T)
  
```

```{r}
genes = (pcgr %>% filter(grepl("TIER [1-4]", TIER, )))$SYMBOL

df = pcgr %>%
filter(grepl("TIER [1-4]", TIER) &
       SYMBOL %in% genes) %>%
select(SYMBOL, VCF_SAMPLE_ID, VARIANT_CLASS) %>%
pivot_wider(names_from = VCF_SAMPLE_ID, 
            values_from = VARIANT_CLASS, 
            id_cols = SYMBOL, 
            values_fill = NA,
            values_fn = list(VARIANT_CLASS = ~paste(., collapse = ";"))) %>%
column_to_rownames('SYMBOL')

# add empty rows for samples not in matrix 
if(length(setdiff(unique(pcgr$VCF_SAMPLE_ID), colnames(df))) > 0){
onco_matrix[setdiff(unique(pcgr$VCF_SAMPLE_ID), colnames(df))] = NA
}

```

# TIER 1 and 2 

## Filter tier 1 and 2 and count by gene 

```{r}
pcgr %>%
filter(TIER %in% c('TIER 1', 'TIER 2')) %>%
count(SYMBOL, sort = T)
```

## Filter tier 1 and 2 and count by gene and genomic change 

```{r}
pcgr %>%
filter(TIER %in% c('TIER 1', 'TIER 2')) %>%
count(SYMBOL, PROTEIN_CHANGE, sort = T)
```

# TIER 3 

## Filter tier 3 and count by gene 

```{r}
pcgr %>%
filter(TIER %in% c('TIER 3')) %>%
count(SYMBOL, sort = T)
```

## Filter tier 3 and count by gene and genomic change 

```{r}
pcgr %>%
filter(TIER %in% c('TIER 3')) %>%
count(SYMBOL, PROTEIN_CHANGE, sort = T)
```

# TIER 4

## Filter tier 4 and count by gene 

```{r}
pcgr %>%
filter(TIER %in% c('TIER 4')) %>%
count(SYMBOL, sort = T)
```

## Filter tier 4 and count by gene and genomic change 

```{r}
pcgr %>%
filter(TIER %in% c('TIER 4')) %>%
count(SYMBOL, PROTEIN_CHANGE, sort = T)
```

# Some recurrent gene families 

## SOX

```{r}
pcgr %>%
filter(grepl('SOX', SYMBOL))
```
## LEF

```{r}
pcgr %>%
filter(grepl('LEF', SYMBOL))
```
## TCF

```{r}
pcgr %>%
filter(grepl('TCF', SYMBOL))
```

## FGF

```{r}
pcgr %>%
filter(grepl('FGF', SYMBOL))
```

## FOX

```{r}
pcgr %>%
filter(grepl('FOX', SYMBOL))
```

## KDM

```{r}
pcgr %>%
filter(grepl('KDM', SYMBOL))
```

# Oncoprint 

```{r}
col = c("SNV" = "#D55E00", 
        "insertion" = "#15587A", 
        "deletion" = "#2A9134", 
        "substitution" = 'red')

alter_fun = list(
  background = function(x, y, w, h) {
    grid.rect(x, y, w-unit(2, "pt"), h-unit(2, "pt"), 
              gp = gpar(fill = "#DEE3E3", col = NA))
  },
  
  SNV = function(x, y, w, h) {
    grid.rect(x, y, w-unit(2, "pt"), h-unit(2, "pt"), 
              gp = gpar(fill = col["SNV"], col = NA))
  },
  
  insertion = function(x, y, w, h) {
    grid.rect(x, y, w-unit(2, "pt"), h*0.5, 
              gp = gpar(fill = col["insertion"], col = NA))
  },
  
  deletion = function(x, y, w, h) {
    grid.rect(x, y, w-unit(2, "pt"), h*0.5, 
              gp = gpar(fill = col["deletion"], col = NA))
  },
  substitution = function(x, y, w, h) {
    grid.rect(x, y, w-unit(2, "pt"), h*0.5, 
              gp = gpar(fill = col["substitution"], col = NA))
  }
)
```

## TIER 1, 2 and 3 full 

```{r}
genes = (pcgr %>% filter(TIER %in% c('TIER 1', 'TIER 2', 'TIER 3')))$SYMBOL
```

```{r}
onco_matrix = pcgr %>%
filter(TIER %in% c('TIER 1', 'TIER 2', 'TIER 3') &
       SYMBOL %in% genes) %>%
select(SYMBOL, VCF_SAMPLE_ID, VARIANT_CLASS) %>%
pivot_wider(names_from = VCF_SAMPLE_ID, 
            values_from = VARIANT_CLASS, 
            id_cols = SYMBOL, 
            values_fill = NA,
            values_fn = list(VARIANT_CLASS = ~paste(., collapse = ";"))) %>%
column_to_rownames('SYMBOL')

# add empty rows for samples not in matrix 
if(length(setdiff(unique(pcgr$VCF_SAMPLE_ID), colnames(onco_matrix))) > 0){
onco_matrix[setdiff(unique(pcgr$VCF_SAMPLE_ID), colnames(onco_matrix))] = NA
}
```

```{r}
meta = meta[match(names(onco_matrix), meta$ID_triceps_signature_profyle),] 

ha = HeatmapAnnotation(
  annotation_name_gp= gpar(fontsize = 15),
  EZHIP = meta$EZHIP_Status,
  Gender = meta$Sex,
  col = list(Gender = c("F" = "#413C58" , "M" = "#F18805"),
             EZHIP = c('positive' = "#C12623", "negative" = "#386FA4"))
 )
```

```{r}
op_123_full = oncoPrint(onco_matrix,
               alter_fun = alter_fun,
               bottom_annotation = ha,
               col = col, 
               pct_side = "right", 
               row_names_side = "left", 
               remove_empty_columns = F, 
               show_column_names = T, 
               right_annotation=NULL, 
               top_annotation = NULL, 
               row_names_gp = gpar(fontsize = 20),
               column_names_gp = gpar(fontsize = 20), 
               heatmap_height = unit(70, "cm"), 
               heatmap_width = unit(20, "cm"))
```


## TIER 1, 2 and 3 

```{r}
genes = (pcgr %>%
filter(TIER %in% c('TIER 1', 'TIER 2', 'TIER 3') & AF_TUMOR > 0.01) %>% 
distinct(SYMBOL, VCF_SAMPLE_ID) %>%
count(SYMBOL) %>%
filter(n > 1))$SYMBOL

genes = unique(c(genes, (pcgr %>% filter(TIER %in% c('TIER 1', 'TIER 2')))$SYMBOL))
```

```{r}
onco_matrix = pcgr %>%
filter(TIER %in% c('TIER 1', 'TIER 2', 'TIER 3') &
       SYMBOL %in% genes) %>%
select(SYMBOL, VCF_SAMPLE_ID, VARIANT_CLASS) %>%
pivot_wider(names_from = VCF_SAMPLE_ID, 
            values_from = VARIANT_CLASS, 
            id_cols = SYMBOL, 
            values_fill = NA,
            values_fn = list(VARIANT_CLASS = ~paste(., collapse = ";"))) %>%
column_to_rownames('SYMBOL')

# add empty rows for samples not in matrix 
if(length(setdiff(unique(pcgr$VCF_SAMPLE_ID), colnames(onco_matrix))) > 0){
onco_matrix[setdiff(unique(pcgr$VCF_SAMPLE_ID), colnames(onco_matrix))] = NA
}
```

```{r}
meta = meta[match(names(onco_matrix), meta$ID_triceps_signature_profyle),] 

ha = HeatmapAnnotation(
  annotation_name_gp= gpar(fontsize = 15),
  EZHIP = meta$EZHIP_Status,
  Gender = meta$Sex,
  col = list(Gender = c("F" = "#413C58" , "M" = "#F18805"),
             EZHIP = c('positive' = "#C12623", "negative" = "#386FA4"))
 )
```

```{r}
op_123 = oncoPrint(onco_matrix,
               alter_fun = alter_fun, 
               bottom_annotation = ha,
               col = col, 
               pct_side = "right", 
               row_names_side = "left", 
               remove_empty_columns = F, 
               show_column_names = T, 
               right_annotation=NULL, 
               top_annotation = NULL, 
               row_names_gp = gpar(fontsize = 20),
               column_names_gp = gpar(fontsize = 20), 
               heatmap_height = unit(25, "cm"), 
               heatmap_width = unit(20, "cm"))
```

## SOX  

```{r}
onco_matrix = pcgr %>%
filter(TIER %in% c('TIER 1', 'TIER 2', 'TIER 3', 'TIER 4') &
       (grepl('SOX', SYMBOL) | grepl('LEF', SYMBOL) | grepl('TCF', SYMBOL) )) %>%
select(SYMBOL, VCF_SAMPLE_ID, VARIANT_CLASS) %>%
pivot_wider(names_from = VCF_SAMPLE_ID, 
            values_from = VARIANT_CLASS, 
            id_cols = SYMBOL, 
            values_fill = NA,
            values_fn = list(VARIANT_CLASS = ~paste(., collapse = ";"))) %>%
column_to_rownames('SYMBOL')

# add empty rows for samples not in matrix 
if(length(setdiff(unique(pcgr$VCF_SAMPLE_ID), colnames(onco_matrix))) > 0){
onco_matrix[setdiff(unique(pcgr$VCF_SAMPLE_ID), colnames(onco_matrix))] = NA
}
```

```{r}
meta = meta[match(names(onco_matrix), meta$ID_triceps_signature_profyle),] 

ha = HeatmapAnnotation(
  annotation_name_gp= gpar(fontsize = 15),
  EZHIP = meta$EZHIP_Status,
  Gender = meta$Sex,
  col = list(Gender = c("F" = "#413C58" , "M" = "#F18805"),
             EZHIP = c('positive' = "#C12623", "negative" = "#386FA4"))
 )
```

```{r}
op_SOX = oncoPrint(onco_matrix,
               alter_fun = alter_fun, 
               bottom_annotation = ha,
               col = col, 
               pct_side = "right", 
               row_names_side = "left", 
               remove_empty_columns = F, 
               show_column_names = T, 
               right_annotation=NULL, 
               top_annotation = NULL, 
               row_names_gp = gpar(fontsize = 20),
               column_names_gp = gpar(fontsize = 20), 
               heatmap_height = unit(20, "cm"), 
               heatmap_width = unit(20, "cm"))
```

## FOX  

```{r}
onco_matrix = pcgr %>%
filter(TIER %in% c('TIER 1', 'TIER 2', 'TIER 3', 'TIER 4') &
       (grepl('FOX', SYMBOL))) %>%
select(SYMBOL, VCF_SAMPLE_ID, VARIANT_CLASS) %>%
pivot_wider(names_from = VCF_SAMPLE_ID, 
            values_from = VARIANT_CLASS, 
            id_cols = SYMBOL, 
            values_fill = NA,
            values_fn = list(VARIANT_CLASS = ~paste(., collapse = ";"))) %>%
column_to_rownames('SYMBOL')

# add empty rows for samples not in matrix 
if(length(setdiff(unique(pcgr$VCF_SAMPLE_ID), colnames(onco_matrix))) > 0){
onco_matrix[setdiff(unique(pcgr$VCF_SAMPLE_ID), colnames(onco_matrix))] = NA
}
```

```{r}
meta = meta[match(names(onco_matrix), meta$ID_triceps_signature_profyle),] 

ha = HeatmapAnnotation(
  annotation_name_gp= gpar(fontsize = 15),
  EZHIP = meta$EZHIP_Status,
  Gender = meta$Sex,
  col = list(Gender = c("F" = "#413C58" , "M" = "#F18805"),
             EZHIP = c('positive' = "#C12623", "negative" = "#386FA4"))
 )
```

```{r}
op_FOX = oncoPrint(onco_matrix,
               alter_fun = alter_fun, 
               bottom_annotation = ha, 
               col = col, 
               pct_side = "right", 
               row_names_side = "left", 
               remove_empty_columns = F, 
               show_column_names = T, 
               right_annotation=NULL, 
               top_annotation = NULL, 
               row_names_gp = gpar(fontsize = 20),
               column_names_gp = gpar(fontsize = 20), 
               heatmap_height = unit(20, "cm"), 
               heatmap_width = unit(20, "cm"))
```

## Meth

```{r}
onco_matrix = pcgr %>%
filter(TIER %in% c('TIER 1', 'TIER 2', 'TIER 3', 'TIER 4') &
       (grepl('KDM6', SYMBOL))) %>%
select(SYMBOL, VCF_SAMPLE_ID, VARIANT_CLASS) %>%
pivot_wider(names_from = VCF_SAMPLE_ID, 
            values_from = VARIANT_CLASS, 
            id_cols = SYMBOL, 
            values_fill = NA,
            values_fn = list(VARIANT_CLASS = ~paste(., collapse = ";"))) %>%
column_to_rownames('SYMBOL')

# add empty rows for samples not in matrix 
if(length(setdiff(unique(pcgr$VCF_SAMPLE_ID), colnames(onco_matrix))) > 0){
onco_matrix[setdiff(unique(pcgr$VCF_SAMPLE_ID), colnames(onco_matrix))] = NA
}
```

```{r}
meta = meta[match(names(onco_matrix), meta$ID_triceps_signature_profyle),] 

ha = HeatmapAnnotation(
  annotation_name_gp= gpar(fontsize = 15),
  EZHIP = meta$EZHIP_Status,
  Gender = meta$Sex,
  col = list(Gender = c("F" = "#413C58" , "M" = "#F18805"),
             EZHIP = c('positive' = "#C12623", "negative" = "#386FA4"))
 )
```

```{r}
op_meth = oncoPrint(onco_matrix,
               alter_fun = alter_fun, 
               bottom_annotation = ha,
               col = col, 
               pct_side = "right", 
               row_names_side = "left", 
               remove_empty_columns = F, 
               show_column_names = T, 
               right_annotation=NULL, 
               top_annotation = NULL, 
               row_names_gp = gpar(fontsize = 20),
               column_names_gp = gpar(fontsize = 20), 
               heatmap_height = unit(20, "cm"), 
               heatmap_width = unit(20, "cm"))
```

## FGF  

```{r}
onco_matrix = pcgr %>%
filter(TIER %in% c('TIER 1', 'TIER 2', 'TIER 3', 'TIER 4') &
       (grepl('FGF', SYMBOL))) %>%
select(SYMBOL, VCF_SAMPLE_ID, VARIANT_CLASS) %>%
pivot_wider(names_from = VCF_SAMPLE_ID, 
            values_from = VARIANT_CLASS, 
            id_cols = SYMBOL, 
            values_fill = NA,
            values_fn = list(VARIANT_CLASS = ~paste(., collapse = ";"))) %>%
column_to_rownames('SYMBOL')

# add empty rows for samples not in matrix 
if(length(setdiff(unique(pcgr$VCF_SAMPLE_ID), colnames(onco_matrix))) > 0){
onco_matrix[setdiff(unique(pcgr$VCF_SAMPLE_ID), colnames(onco_matrix))] = NA
}
```

```{r}
meta = meta[match(names(onco_matrix), meta$ID_triceps_signature_profyle),] 

ha = HeatmapAnnotation(
  annotation_name_gp= gpar(fontsize = 15),
  EZHIP = meta$EZHIP_Status,
  Gender = meta$Sex,
  col = list(Gender = c("F" = "#413C58" , "M" = "#F18805"),
             EZHIP = c('positive' = "#C12623", "negative" = "#386FA4"))
 )
```

```{r}
op_FGF = oncoPrint(onco_matrix,
               alter_fun = alter_fun, 
               bottom_annotation = ha,
               col = col, 
               pct_side = "right", 
               row_names_side = "left", 
               remove_empty_columns = F, 
               show_column_names = T, 
               right_annotation=NULL, 
               top_annotation = NULL, 
               row_names_gp = gpar(fontsize = 20),
               column_names_gp = gpar(fontsize = 20), 
               heatmap_height = unit(20, "cm"), 
               heatmap_width = unit(20, "cm"))
```


## Histone  

```{r}
hist_genes = msigdbr(species = "human", category = 'C5') %>%          
                      dplyr::select(c("gs_name", "gene_symbol", "human_ensembl_gene")) %>%
                      dplyr::filter(grepl('HISTONE', gs_name))

hist_genes = hist_genes[grepl('ASH|DOT|NSD|KDM|KMT|SET|KAT|HDAC', hist_genes$gene_symbol),]
```


```{r}
onco_matrix = pcgr %>%
filter(TIER %in% c('TIER 1', 'TIER 2', 'TIER 3', 'TIER 4') & (ENSEMBL_GENE_ID %in% c(hist_genes$human_ensembl_gene)) | grepl('SETBP1|HIST|CXorf67|EP300', SYMBOL)) %>%
  filter(CODING_STATUS == 'coding') %>%
select(SYMBOL, VCF_SAMPLE_ID, VARIANT_CLASS) %>% 
pivot_wider(names_from = VCF_SAMPLE_ID, 
            values_from = VARIANT_CLASS, 
            id_cols = SYMBOL, 
            values_fill = NA,
            values_fn = list(VARIANT_CLASS = ~paste(., collapse = ";"))) %>%
column_to_rownames('SYMBOL')

# add empty rows for samples not in matrix 
if(length(setdiff(unique(pcgr$VCF_SAMPLE_ID), colnames(onco_matrix))) > 0){
onco_matrix[setdiff(unique(pcgr$VCF_SAMPLE_ID), colnames(onco_matrix))] = NA
}
```

```{r}
meta = meta[match(names(onco_matrix), meta$ID_triceps_signature_profyle),] 

ha = HeatmapAnnotation(
  annotation_name_gp= gpar(fontsize = 15),
  EZHIP = meta$EZHIP_Status,
  Gender = meta$Sex,
  col = list(Gender = c("F" = "#413C58" , "M" = "#F18805"),
             EZHIP = c('positive' = "#C12623", "negative" = "#386FA4"))
 )
```

```{r}
op_histone = oncoPrint(onco_matrix,
               alter_fun = alter_fun, 
               bottom_annotation = ha,
               col = col, 
               pct_side = "right", 
               row_names_side = "left", 
               remove_empty_columns = F, 
               show_column_names = T, 
               right_annotation=NULL, 
               top_annotation = NULL, 
               row_names_gp = gpar(fontsize = 20),
               column_names_gp = gpar(fontsize = 20), 
               heatmap_height = unit(20, "cm"), 
               heatmap_width = unit(20, "cm"))
```

## Save plots 

```{r}
pdf('/lustre06/project/6004736/alvann/from_narval/220318_EZHIP/WES/clinical_seq_patients/out/02_oncoprint/oncoprint.pdf', width = 20, height = 30)
draw(op_123_full)
draw(op_123)
draw(op_histone)
dev.off()
```


