---
title: "Oncoprint"
output: 
  html_document:
    df_print: paged
    theme: flatly
    number_sections: yes
    toc: yes
    toc_float: yes
    code_folding: hide
---

```{r}
library(tidyverse)
library(ComplexHeatmap)
library(msigdbr)
```

```{r}
PATH_GERM = "/lustre06/project/6004736/alvann/from_narval/220318_EZHIP/integrated/221106_oncoprint/out/01_concat/germline.cpsr.tiers.all.samples.tsv"
PATH_SOM = "/lustre06/project/6004736/alvann/from_narval/220318_EZHIP/integrated/221106_oncoprint/out/01_concat/somatic.pcgr.tiers.all.samples.tsv"
PATH_META_SAMPLE = "/lustre06/project/6004736/alvann/from_narval/220318_EZHIP/files/sample_meta.tsv"
PATH_META_PATIENT = "/lustre06/project/6004736/alvann/from_narval/220318_EZHIP/files/patient_meta.tsv"
```

```{r}
cpsr = data.table::fread(PATH_GERM) %>%
       filter(CLINVAR_CLASSIFICATION %in% c('Pathogenic', 'Likely-Pathogenic')) %>%
       separate(VCF_SAMPLE_ID, into = c('PATIENT', NA, 'RELAPSE'), sep = '_', remove = F) %>%
       mutate(RELAPSE = ifelse(RELAPSE == 'T', '0T', RELAPSE), 
              RELAPSE = ifelse(is.na(RELAPSE), '0T', RELAPSE))
```

```{r}
pcgr = data.table::fread(PATH_SOM) %>%
       filter(CODING_STATUS == 'coding') %>%
       filter(!(is.na(AF_CONTROL) & AF_TUMOR < 0.1)) %>%
       filter(DP_TUMOR >=10) %>%
       separate(VCF_SAMPLE_ID, into = c('PATIENT', NA, 'RELAPSE'), sep = '_', remove = F) %>%
       mutate(RELAPSE = ifelse(RELAPSE == 'T', '0T', RELAPSE), 
              RELAPSE = ifelse(is.na(RELAPSE), '0T', RELAPSE))
```

```{r}
keep = (pcgr %>%
  select(VCF_SAMPLE_ID, PATIENT, RELAPSE) %>%
  distinct() %>%
  arrange(PATIENT, RELAPSE) %>%
  group_by(PATIENT) %>%
  top_n(wt = RELAPSE, n = -1))$VCF_SAMPLE_ID
  
```

```{r}
pcgr = pcgr %>%
       filter(VCF_SAMPLE_ID %in% keep)

cpsr = cpsr %>%
       filter(VCF_SAMPLE_ID %in% keep)
```


```{r}
meta_pat = data.table::fread(PATH_META_PATIENT) %>% arrange(Patient_N)
meta_samp = data.table::fread(PATH_META_SAMPLE)
```

```{r}
meta = full_join(meta_samp, meta_pat, by = 'Patient_N') %>%
       filter(!is.na(Patient_N)) %>%
       mutate(Response = tolower(Response), 
              Primary_location = tolower(Primary_location)) %>%
       mutate(Sample_oncoprint = ifelse(is.na(ID_triceps_signature_profyle), Sample, ID_triceps_signature_profyle)) 

meta = meta %>%
       filter(Sample_oncoprint %in% intersect(pcgr$PATIENT, meta$Sample_oncoprint))
```

```{r}
pcgr = pcgr %>%
  left_join(meta %>% select(Sample_oncoprint, Patient_N), by = c('PATIENT' = 'Sample_oncoprint')) %>%
  filter(PATIENT %in% intersect(pcgr$PATIENT, meta$Sample_oncoprint))

cpsr = cpsr %>%
  left_join(meta %>% select(Sample_oncoprint, Patient_N), by = c('PATIENT' = 'Sample_oncoprint')) %>%
  filter(PATIENT %in% intersect(pcgr$PATIENT, meta$Sample_oncoprint))
```

```{r}
onco_matrix = bind_rows(pcgr, cpsr %>% mutate(VARIANT_CLASS = paste(VARIANT_CLASS, 'germline', sep = ';'))) %>%
  filter(grepl("TIER [1-3]", TIER) | SYMBOL %in% c('H3F3A', 'H3F3B') | grepl('germline', VARIANT_CLASS)) %>%
select(SYMBOL, Patient_N, VARIANT_CLASS) %>%
pivot_wider(names_from = Patient_N, 
            values_from = VARIANT_CLASS, 
            id_cols = SYMBOL, 
            values_fill = NA,
            values_fn = list(VARIANT_CLASS = ~paste(., collapse = ";"))) %>%
column_to_rownames('SYMBOL')

# add empty rows for samples not in matrix 
if(length(setdiff(unique(pcgr$Patient_N), colnames(onco_matrix))) > 0){
onco_matrix[as.character(setdiff(unique(pcgr$Patient_N), colnames(onco_matrix)))] = NA
}
```

```{r}
data.table::fwrite(onco_matrix, "/lustre06/project/6004736/alvann/from_narval/220318_EZHIP/integrated/221106_oncoprint/out/02_oncoprint_snv/oncomat_snv.tsv", row.names = T) 
```

```{r}

```


# All TIERS 

## Count by gene 

```{r}
pcgr %>%
  filter(grepl('TIER', pcgr$TIER)) %>%
  count(SYMBOL, sort = T)
```

## Count by gene and genomic change 

```{r}
pcgr %>% 
  filter(grepl('TIER', pcgr$TIER)) %>%
  count(SYMBOL, PROTEIN_CHANGE, sort = T)
```

```{r}
genes = (pcgr %>% filter(grepl("TIER [1-4]", TIER, )))$SYMBOL

df = pcgr %>%
filter(grepl("TIER [1-4]", TIER) &
       SYMBOL %in% genes) %>%
select(SYMBOL, VCF_SAMPLE_ID, VARIANT_CLASS) %>%
pivot_wider(names_from = VCF_SAMPLE_ID, 
            values_from = VARIANT_CLASS, 
            id_cols = SYMBOL, 
            values_fill = NA,
            values_fn = list(VARIANT_CLASS = ~paste(., collapse = ";"))) %>%
column_to_rownames('SYMBOL')

# add empty rows for samples not in matrix 
if(length(setdiff(unique(pcgr$VCF_SAMPLE_ID), colnames(df))) > 0){
onco_matrix[setdiff(unique(pcgr$VCF_SAMPLE_ID), colnames(df))] = NA
}

```

# TIER 1 and 2 

## Filter tier 1 and 2 and count by gene 

```{r}
pcgr %>%
filter(TIER %in% c('TIER 1', 'TIER 2')) %>%
count(SYMBOL, sort = T)
```

## Filter tier 1 and 2 and count by gene and genomic change 

```{r}
pcgr %>%
filter(TIER %in% c('TIER 1', 'TIER 2')) %>%
count(SYMBOL, PROTEIN_CHANGE, sort = T)
```

# TIER 3 

## Filter tier 3 and count by gene 

```{r}
pcgr %>%
filter(TIER %in% c('TIER 3')) %>%
count(SYMBOL, sort = T)
```

## Filter tier 3 and count by gene and genomic change 

```{r}
pcgr %>%
filter(TIER %in% c('TIER 3')) %>%
count(SYMBOL, PROTEIN_CHANGE, sort = T)
```

# TIER 4

## Filter tier 4 and count by gene 

```{r}
pcgr %>%
filter(TIER %in% c('TIER 4')) %>%
count(SYMBOL, sort = T)
```

## Filter tier 4 and count by gene and genomic change 

```{r}
pcgr %>%
filter(TIER %in% c('TIER 4')) %>%
count(SYMBOL, PROTEIN_CHANGE, sort = T)
```

# Some recurrent gene families 

## SOX

```{r}
pcgr %>%
filter(grepl('SOX', SYMBOL))
```
## LEF

```{r}
pcgr %>%
filter(grepl('LEF', SYMBOL))
```
## TCF

```{r}
pcgr %>%
filter(grepl('TCF', SYMBOL))
```

## FGF

```{r}
pcgr %>%
filter(grepl('FGF', SYMBOL))
```

```{r}
pcgr %>%
filter(grepl('^IGF', SYMBOL))
```

## FOX

```{r}
pcgr %>%
filter(grepl('FOX', SYMBOL))
```

## KDM

```{r}
pcgr %>%
filter(grepl('KDM', SYMBOL))
```

## KTM

```{r}
pcgr %>%
filter(grepl('^KMT', SYMBOL))
```

# Oncoprint 

```{r}
col = c("SNV" = "#D55E00", 
        "insertion" = "#15587A", 
        "deletion" = "#2A9134", 
        "substitution" = 'red')

alter_fun = list(
  background = function(x, y, w, h) {
    grid.rect(x, y, w-unit(2, "pt"), h-unit(2, "pt"), 
              gp = gpar(fill = "#DEE3E3", col = NA))
  },
  
  SNV = function(x, y, w, h) {
    grid.rect(x, y, w-unit(2, "pt"), h-unit(2, "pt"), 
              gp = gpar(fill = col["SNV"], col = NA))
  },
  
  insertion = function(x, y, w, h) {
    grid.rect(x, y, w-unit(2, "pt"), h*0.5, 
              gp = gpar(fill = col["insertion"], col = NA))
  },
  
  deletion = function(x, y, w, h) {
    grid.rect(x, y, w-unit(2, "pt"), h*0.5, 
              gp = gpar(fill = col["deletion"], col = NA))
  },
  substitution = function(x, y, w, h) {
    grid.rect(x, y, w-unit(2, "pt"), h*0.5, 
              gp = gpar(fill = col["substitution"], col = NA))
  },
  germline = function(x, y, w, h) 
        grid.segments(x - w*0.5, y - h*0.5, x + w*0.5, y + h*0.5, gp = gpar(lwd = 2))
)
```

## TIER 1, 2 and 3 full 

```{r}
genes = (pcgr %>% filter(TIER %in% c('TIER 1', 'TIER 2', 'TIER 3')))$SYMBOL
```

```{r}
onco_matrix = pcgr %>%
filter(TIER %in% c('TIER 1', 'TIER 2', 'TIER 3') &
       SYMBOL %in% genes) %>%
select(SYMBOL, Patient_N, VARIANT_CLASS) %>%
pivot_wider(names_from = Patient_N, 
            values_from = VARIANT_CLASS, 
            id_cols = SYMBOL, 
            values_fill = NA,
            values_fn = list(VARIANT_CLASS = ~paste(., collapse = ";"))) %>%
column_to_rownames('SYMBOL')

# add empty rows for samples not in matrix 
if(length(setdiff(unique(pcgr$Patient_N), colnames(onco_matrix))) > 0){
onco_matrix[as.character(setdiff(unique(pcgr$Patient_N), colnames(onco_matrix)))] = NA
}
```

```{r}
meta = meta[match(names(onco_matrix), meta$Patient_N),] 

ha = HeatmapAnnotation(
  annotation_name_gp= gpar(fontsize = 15),
  EZHIP = meta$EZHIP_Status,
  Gender = meta$Sex,
  col = list(Gender = c("F" = "#413C58" , "M" = "#F18805"),
             EZHIP = c('positive' = "#C12623", "negative" = "#386FA4"))
 )
```

```{r}
op_123_full = oncoPrint(onco_matrix,
               alter_fun = alter_fun,
               bottom_annotation = ha,
               col = col, 
               pct_side = "right", 
               row_names_side = "left", 
               remove_empty_columns = F, 
               show_column_names = T, 
               right_annotation=NULL, 
               top_annotation = NULL, 
               row_names_gp = gpar(fontsize = 20),
               column_names_gp = gpar(fontsize = 20), 
               width = ncol(onco_matrix)*unit(10, "mm"), 
               height = nrow(onco_matrix)*unit(10, "mm"))
```


## TIER 1, 2 and 3 

```{r}
genes = (pcgr %>%
filter(TIER %in% c('TIER 1', 'TIER 2', 'TIER 3')) %>% 
distinct(SYMBOL, VCF_SAMPLE_ID) %>%
count(SYMBOL) %>%
filter(n >=2))$SYMBOL

genes = unique(c(genes, (pcgr %>% filter(TIER %in% c('TIER 1', 'TIER 2')))$SYMBOL))
```

```{r}
onco_matrix = bind_rows(pcgr, cpsr %>% mutate(VARIANT_CLASS = paste(VARIANT_CLASS, 'germline', sep = ';'))) %>%
filter((TIER %in% c('TIER 1', 'TIER 2', 'TIER 3') | grepl('germline', VARIANT_CLASS)) &
       SYMBOL %in% genes) %>%
select(SYMBOL, PATIENT, Patient_N, VARIANT_CLASS) %>%
pivot_wider(names_from = Patient_N, 
            values_from = VARIANT_CLASS, 
            id_cols = SYMBOL, 
            values_fill = NA,
            values_fn = list(VARIANT_CLASS = ~paste(., collapse = ";"))) %>%
column_to_rownames('SYMBOL')

# add empty rows for samples not in matrix 
if(length(setdiff(unique(pcgr$Patient_N), colnames(onco_matrix))) > 0){
onco_matrix[as.character(setdiff(unique(pcgr$Patient_N), colnames(onco_matrix)))] = NA
}
```

```{r}
meta = meta[match(names(onco_matrix), meta$Patient_N),] 

ha = HeatmapAnnotation(
  annotation_name_gp= gpar(fontsize = 15),
  EZHIP = meta$EZHIP_Status,
  Gender = meta$Sex,
  col = list(Gender = c("F" = "#413C58" , "M" = "#F18805"),
             EZHIP = c('positive' = "#C12623", "negative" = "#386FA4"))
 )
```

```{r}
op_123 = oncoPrint(onco_matrix,
               alter_fun = alter_fun, 
               bottom_annotation = ha,
               col = col, 
               pct_side = "right", 
               row_names_side = "left", 
               remove_empty_columns = F, 
               show_column_names = T, 
               right_annotation=NULL, 
               top_annotation = NULL, 
               row_names_gp = gpar(fontsize = 20),
               column_names_gp = gpar(fontsize = 20), 
               width = ncol(onco_matrix)*unit(10, "mm"), 
               height = nrow(onco_matrix)*unit(10, "mm"))
```


## Histone  

```{r}
hist_genes = msigdbr(species = "human", category = 'C5') %>%          
                      dplyr::select(c("gs_name", "gene_symbol", "human_ensembl_gene")) %>%
                      dplyr::filter(grepl('HISTONE', gs_name))

hist_genes = hist_genes[grepl('ASH|DOT|NSD|KDM|KMT|SET|KAT|HDAC', hist_genes$gene_symbol),]
```


```{r}
onco_matrix = pcgr %>%
filter(TIER %in% c('TIER 1', 'TIER 2', 'TIER 3', 'TIER 4') & (ENSEMBL_GENE_ID %in% c(hist_genes$human_ensembl_gene)) | grepl('SETBP1|HIST|CXorf67|EP300|H3F3', SYMBOL)) %>%
  filter(CODING_STATUS == 'coding') %>%
select(SYMBOL, Patient_N, VARIANT_CLASS) %>% 
pivot_wider(names_from = Patient_N, 
            values_from = VARIANT_CLASS, 
            id_cols = SYMBOL, 
            values_fill = NA,
            values_fn = list(VARIANT_CLASS = ~paste(., collapse = ";"))) %>%
column_to_rownames('SYMBOL')

# add empty rows for samples not in matrix 
if(length(setdiff(unique(pcgr$Patient_N), colnames(onco_matrix))) > 0){
onco_matrix[as.character(setdiff(unique(pcgr$Patient_N), colnames(onco_matrix)))] = NA
}
```

```{r}
meta = meta[match(names(onco_matrix), meta$Patient_N),] 

ha = HeatmapAnnotation(
  annotation_name_gp= gpar(fontsize = 15),
  EZHIP = meta$EZHIP_Status,
  Gender = meta$Sex,
  col = list(Gender = c("F" = "#413C58" , "M" = "#F18805"),
             EZHIP = c('positive' = "#C12623", "negative" = "#386FA4"))
 )
```

```{r}
op_histone = oncoPrint(onco_matrix,
               alter_fun = alter_fun, 
               bottom_annotation = ha,
               col = col, 
               pct_side = "right", 
               row_names_side = "left", 
               remove_empty_columns = F, 
               show_column_names = T, 
               right_annotation=NULL, 
               top_annotation = NULL, 
               row_names_gp = gpar(fontsize = 20),
               column_names_gp = gpar(fontsize = 20), 
               width = ncol(onco_matrix)*unit(10, "mm"), 
               height = nrow(onco_matrix)*unit(10, "mm"))
```

## Save plots 

```{r}
pdf('/lustre06/project/6004736/alvann/from_narval/220318_EZHIP/integrated/221106_oncoprint/out/02_oncoprint_snv/oncoprint_snv.pdf', width = 20, height = 30)
draw(op_123_full)
draw(op_123)
draw(op_histone)
dev.off()
```

```{r}
cpsr %>% filter(SYMBOL %in%  c('TP53', 'RB1')) 
```




